The Linear Instability of Dilute Ultrarelativistic \epm\ Pair Beams

### 2016-06-04 oblique.tex
final changes

Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
Taking the limit of Equation (\ref{eq:distribution function beam}) as $\alpha_{\parallel}^{-1}, \alpha_{\perp}^{-1} \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
\begin{equation}\label{eq:dispersion relation reactive}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
\end{equation}
For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{BCP}.

@@

@@

\begin{figure*}
\includegraphics[width=0.9\textwidth]{growthrates.png}
\caption{Oblique growth rate maximized over $\vph$ as a function of
  $v_{bx}$ and $T$ for various $\gamma_b$.  In each the green line
  shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.  In all
  cases, the maximum growth rate occurs when
  $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
\end{figure*}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{oblique_GammaM.eps}
\caption{Oblique kinetic growth rate, maximized over $v_{bx}/v_b$ and
  $\omega/k$, normalized by
  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function
  of $T$ (in units of $m$) and $\gamma_b$.  Note that unlike the beam
  plasma case shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the
  transition between high and low temperature is only marginal,
  constituting a roughly $10\%$ reduction.} \label{fig:OGgen}
\end{figure}

@@

Applying these results to the case of ultrarelativistic beams and given our results from Section \ref{sec:setup}, we find
\begin{eqnarray}
\frac{\Delta v_{\perp}}{v_b} &\sim&  \sqrt{\alpha_{\perp}} = \sqrt{\alpha'_{\perp}} \gamma^{-1},\\
\frac{\Delta v_{\parallel}}{v_b} &\sim& \sqrt{\alpha_{\parallel}} = \sqrt{\alpha'_{\parallel}}\gamma^{-2},
\end{eqnarray}
where $\alpha'_{\perp}$ and $\alpha'_{\parallel}$ encodes the dependence on the COM velocity dispersion.  Hence, the conditions for the reactive regime for the ``oblique'' mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
\begin{equation}\label{eq:generic reactive regime}
 \displaystyle
1 \ll \left\{
 \begin{array}{cl}
\alpha_{\perp}'^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
\alpha_{\parallel}'^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
\end{array}
\right..
\end{equation}

We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:growth-rate-bp}]) and ``oblique'' mode (eq.[\ref{eq:growth-rate-oblique}]).  Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.  Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$  For the beam plasma case, the growth rate is roughly
\begin{equation}
 \Gamma \sim \frac{n_b}{\gamma^3n_t} \left(\frac{c}{\Delta v_{\parallel}}\right)^{2}.
\end{equation}
The bandwidth, $\Delta \omega$, is then greater than the growth rate if 
\begin{equation}\label{eq:bp kinetic regime}
 \frac{\Delta v_{\parallel}}{v_b} \gtrsim \gamma^{-1}\left(\frac{n_b}{n_t}\right)^{1/3},
\end{equation}
which connects with the condition on the reactive beam plasma instability from Equation (\ref{eq:bp reactive regime}).
Similarly for the ``oblique'' mode, the bandwidth, $\Delta \omega$, is then greater than the growth rate if 
\begin{equation}\label{eq:oblique kinetic regime}
\frac{\Delta v_{\perp}}{v_b} \gtrsim \left(\frac{n_b}{\gamma n_t}\right)^{1/3},
\end{equation}
which can similarly compared to the condition on the reactive ``oblique'' mode from Equation (\ref{eq:oblique reactive regime}).

Combining these two kinetic condition and our result again from Section \ref{sec:setup}, we find  
\begin{equation}\label{eq:generic kinetic regime}
1 \ge \left\{
\begin{array}{cl}
\alpha_{\perp}'^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
\alpha_{\parallel}'^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
\end{array}
\right.,
\end{equation}
which in combination with Equation (\ref{eq:generic reactive regime}) denotes the transition between the reactive and kinetic regimes.

@@

Still some progress can be made if we assume that the dominant loss rate is ICC so that we can place a lower limit on the various plasma processes.  In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (Paper I):
\begin{eqnarray}
\frac{\nb}{\nIGM} &\simeq&
\frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \nonumber\\
&\simeq &
3\times10^{-18}
\left(\frac{1+z}{1}\right)^{3\zeta-7}
\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)
\left(\frac{E}{\TeV}\right)
\,\cm^{-3}\,,
\label{eq:nbIC}
\end{eqnarray}
where $\Dpp$ is defined in the Introduction to be the mean free path of a VHEGR, $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and $\zeta$ is.

In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.[\ref{eq:generic reactive regime}]) and kinetic regimes (eq.[\ref{eq:generic kinetic regime}]) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.  Applying these conditions (eqns. [\ref{eq:generic reactive regime}] and [\ref{eq:generic kinetic regime}]) to the ultrarelativistic $e^+e^-$ pair beams of interest, we find for the controlling parameter: 
\begin{eqnarray}
\frac{\gamma_b^{2/3}}{\sqrt{\alpha_{\perp}'}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 7\times 10^{-3} \frac{\gamma_6^{2/3}}{\sqrt{\alpha_{\perp}'}}\left(1+z\right)^{\zeta-7/3}\nonumber\\
&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam oblique condition}
\end{eqnarray}
and
\begin{eqnarray}
\frac{\gamma_b}{\sqrt{\alpha_{\parallel}'}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 0.7 \frac{\gamma_6}{\sqrt{\alpha_{\parallel}'}}\left(1+z\right)^{\zeta-7/3}\nonumber\\
&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam bp condition}
\end{eqnarray}
where $\gamma_6 = \gamma/10^6$. 
We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the ``oblique'' instability always exists in the kinetic regime, but the beam plasma instability can also sit in the reactive regime for sufficiently cold beams $\alpha_{\parallel} \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$.

The discussion above confirmed our earlier estimate in Paper I that the ``oblique'' instability sits in the kinetic regime.  However, our new results suggest that the beam-plasma instability matches the ``oblique'' instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.  However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.  First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.  This is a part of ongoing work and will be presented in a future publication. Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{\alpha_{\parallel}'}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.  Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).  Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{Puchwein+2011}.

@@

The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and ``oblique'' instability.  In this work, we have these linear instabilities as they would apply to the ultrarelativistic pair beams.  Our main findings are:
\begin{itemize}
\item{We calculate how the parallel and perpendicular velocity dispersions of a distribution function responds under Lorentz boosts.  We show that these two velocities dispersions respectively scale like $\gamma^{-2}$ and $\gamma^{-1}$, allowing us to develop a simplified distribution function of the TeV electron-positron beam. We also show that this can be identified with a drifting relativistic Maxwell-Juttner distribution. }
\item{Using this simplified distribution function, we have analytically calculated the beam-plasma and ``oblique'' instabilities in both the reactive and kinetic regime.  We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the ``oblique'' mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.  In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
\item{We have also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and ``oblique'' mode, the condition for transition between the kinetic and reactive regimes for both are different.  In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the ``oblique'' mode.  This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.}
\item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectra of the EBL.  We found that the average energy of these beams range from $E_e/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.  This is due to the rapid cutoff in the EBL above $10\,\eV$. }
\item{Using this average COM energy, we presume that the pair beam can be modelled as a relativistic Maxwellian in the COM frame.   We then demonstrate that the ``oblique'' mode is in the kinetic regime. However, the beam plasma mode is marginally in the reactive regime, so that the use of either regime produces roughly the same growth rate.  Moreover, the inclusion of plasma instability losses on the number density of beam particles may place the beam plasma mode firmly in the kinetic regime.}
\end{itemize}

The results of this paper are a first step toward an understanding of the nonlinear regime of these plasma instabilities for ultrarelativistic \epm\ beams.  In particular, by developing a simple distribution function, whose modes can be analytically calculated, we will develop a quasilinear theory for these modes in a forthcoming publication.  Such a quasilinear calculation will capture the dominant wave-particle interaction and capture the initial nonlinear evolution of this instability.

@@

%\section{Lorentz Dependence of the Distribution Function and Velocity Dispersion}
%\label{sec:lorentz}
%Here we explicitly derive the scaling of the parallel and
%  perpendicular velocity dispersions with the Lorentz factor upon boosting the
%  distribution function to the lab frame. Let us begin with a distribution
%function that is isotropic in the COM frame and depends only on energy.
%Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
%function is  $\fCM(E_{\rm CM}(\ppos_{\rm CM},\pmom_{\rm CM}))$.  When we move to the lab
%(denoted with subscript ``L'') frame, the integral of the distribution function remains invariant,
%i.e., total number, or
%\begin{equation}
% \int \fLAB \dVLAB = \int \fCM \dVCM.
%\end{equation}
%It is well-known that under Lorentz transformations (Landau), 
%\begin{equation}
% \dVLAB = \dVCM,
%\end{equation}
%so therefore,
%\begin{equation}
%  \fLAB(\ppos_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM}),\pmom_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
%\end{equation}
%Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
%\begin{equation}
% \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
%\end{equation}
%We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$. More precisely, the initial inertial frame is the COM frame and the lab frame is moving with velocity $\betaBOOST = -|\betaBOOST |$ with respect to the COM frame. This gives:
%\begin{eqnarray}\label{eq:betabarLAB}
% \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
%&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
%\end{eqnarray}
%Breaking the components of \betabarLAB\ into components parallel and
%perpendicular to the boost, we find: 
%\begin{eqnarray}
%\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
%\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
%\end{eqnarray}
%where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
%For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
%However, this Galilean result no longer holds for relativistic motion.
%Now we consider the dispersion around $\betabar$.  In components, the COM frame is:
%\begin{equation}
% \dbetabarCMi = \int \beta_i^2 \fCM \dVCM.
%\end{equation}
%In the lab frame, it is again useful to break it into components -- the parallel component becomes 
%\begin{eqnarray}
%  \dbetabarLABpara &=& \int (\beta_{\rm L, \parallel}-\betabarLABpara)^2 \fLAB \dVLAB ,\nonumber\\ 
% &=& \int \beta_{\rm L,\parallel}^2 \fLAB \dVLAB - \betabarLABpara^2,\nonumber\\

>>>>

Starting with the dispersion relation (\ref{eq:dispersion relation by parts}), we first consider the instability of a cold plasma beam. 
Taking the limit of Equation (\ref{eq:distribution function beam}) as $k_BT  \rightarrow 0$, for a target plasma $v_0=0$ and a beam plasma $v_0=v_b$\footnote{That is, we set $F_s(\bmath{p})=\delta^3(\bmath{p}-\bmath{p}_{0s})$ where $\bmath{p}_{0s}\equiv \gamma_0 m_e v_0 \hat{\bmath{z}}$ is the momentum associated with $v_0$.}, we find
\begin{equation}\label{eq:dispersion relation reactive}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2}\frac{\gamma^2 k_x^2 + k_z^2}{k_x^2 + k_z^2} = 0.
\end{equation}
For $k_x = 0$, we recover the same beam-plasma instability which was described in the Appendix of \citet{BCP}.

@@

\subsubsection{Sub-maximal instabilities}

Equation (\ref{eq:dispersion relation reactive}) can also be solved numerically in terms of $k_x$ and $k_z$.  Here let us specialize to the case of $k_x = 0$, i.e., the beam-plasma case.  In this case, we have 
\begin{equation}\label{eq:dispersion relation beam-plasma}
 1 - \frac{\omega_{p,t}^2}{\omega^2} - \frac{\omega_{p,b}^2}{\gamma^3(\omega-k_z v_b)^2} = 0,
\end{equation}
which we can numerically solve in terms of $\omega/\omega_{p,t}$, $k_z\lambda_D$, $v_b/c$, and $n_b/n_t$, where $\lambda_D = c/\omega_{p,t}$ is the skin depth.  In Figure \ref{fig:OffResonance} we show the real and imaginary parts for $\omega/\omega_{p,t}$ as a function of $k_z\lambda_D$ for representative case of $v_b/c \approx 1$ and $n_b/n_t = 10^{-3}$ and $\gamma=100$.    For $k_z\lambda_D = 1$, the growth rate reaches it maximum of $\Gamma_{\rm max}$ and the real part of the frequency is $\Re(\omega) = \omega_{p,t}$, which is plasma oscillation frequency.  This wave would exist in the absence of a tenuous beam.  However, as we move away from this frequency toward lower $k_z$, we still find substantial growth, with $\Gamma \approx 0.1 \Gamma_{\rm max}$ as $k_z \approx 0.9\lambda_D^{-1}$.  Interestingly, the real part of the unstable wave has a phase velocity, $v_{\rm ph} = \Re(\omega)/k = c$, which is still in resonance with the beam.  This unstable wave does not exist in the absence of the destabilizing beam and does not correspond to any wave in the unperturbed system.

In a continuous system, these waves do not matter compare to the unstable mode at $k\lambda_D = 1$.   However, for discrete numerical systems, which do not sufficiently resolve the most unstable modes, these sub maximal modes drive the growth of the instability of numerically calculated beam-plasma systems, which may lead to an incorrect nonlinear state compare to the physical system.

\begin{figure}
\includegraphics[width=0.5\textwidth]{off_res.pdf}
\caption{Beam-plasma growth rate (solid line) and unstable wave frequency (dashed-line) as a function of
  $k_z$ for $\gamma_b = 100$ and $n_b/n_t=10^{-3}$. \label{fig:OffResonance}}
\end{figure}

@@

\begin{figure}
\includegraphics[width=0.5\textwidth]{pp1.pdf}
\caption{Oblique growth rate maximized over $\vph$ as a function of
  $v_{bx}$ and $T$ for various $\gamma_b$.  In each the green line
  shows the chord along which $v_{bx}/v_b = 1/\gamma_b$.  In all
  cases, the maximum growth rate occurs when
  $v_{bx}/v_b\gg1/\gamma_b$.}\label{fig:ObliqueGMZ}
\end{figure}

@@

\begin{figure*}
\includegraphics[width=0.5\textwidth]{pp2.pdf}
\includegraphics[width=0.5\textwidth]{pp3.pdf}

%\plottwo{pp2.pdf}{pp3.pdf}
\caption{Oblique kinetic growth rate (left) and beam-plasma growth rate (right), maximized over $\theta$ and normalized by
  $\Gamma_0\equiv\omega_{Pt}(n_b/n_t)\gamma_b m v_b^2/T$ as a function
  of $k_BT/m_ec^2$ and $\gamma_b$.  Note that unlike the beam
  plasma case shown in Figure \ref{fig:Ggen}, at high $\gamma_b$ the
  transition between high and low temperature is only marginal,
  constituting a roughly $10\%$ reduction.} \label{fig:OGgen}
\end{figure*}

@@

In Appendix \ref{sec:lorentz}, we estimate how the parallel and perpendicular velocity spreads scale in a Maxwell-Junter distribution.  The relevant results are: 
\begin{equation}
\frac{\Delta v_{\perp}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^2 m_e c^2 },
\end{equation}
and 
\begin{equation}
\frac{\Delta v_{\parallel}^2}{c^2} \approx \frac {k_B T}{\gamma_{\rm B}^4 m_e c^2 },
\end{equation}
where $T$ is measured in the COM frame of the beam.

Applying these results, we find
\begin{eqnarray}
\frac{\Delta v_{\perp}}{v_b} &\sim&  \gamma_b^{-1}\sqrt{\frac{k_BT}{m_ec^2}},\\
\frac{\Delta v_{\parallel}}{v_b} &\sim& \gamma_b^{-2}\sqrt{\frac{k_BT}{m_ec^2}}.
\end{eqnarray}
Hence, the conditions for the reactive regime for the ``oblique'' mode (eq.[\ref{eq:oblique reactive regime}]) and beam-plasma mode (eq.[\ref{eq:bp reactive regime}]) can be reduced to 
\begin{equation}\label{eq:generic reactive regime}
 \displaystyle
1 \ll \left\{
 \begin{array}{cl}
 ({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
\end{array}
\right..
\end{equation}

We now proceed to study the range of validity for the kinetic growth rate for the beam plasma mode (eq.[\ref{eq:bp kinetic growth rate}]) and ``oblique'' mode (eq.[\ref{eq:kinetic growth rate}]).  Following the argument of \cite{Boyd}, the growth occurs over a range where the distribution function is positive or $v_b - \Delta v < \omega/k < v_b$.  Hence the bandwidth over which the distribution powers grows is $\Delta \omega \sim k \Delta v$  For the beam plasma case, the growth rate is roughly
\begin{equation}
 \Gamma \sim \frac{n_b}{\gamma^3n_t} \left(\frac{c}{\Delta v_{\parallel}}\right)^{2}.
\end{equation}
The bandwidth, $\Delta \omega$, is then greater than the growth rate if 
\begin{equation}\label{eq:bp kinetic regime}
 \frac{\Delta v_{\parallel}}{v_b} \gtrsim \gamma^{-1}\left(\frac{n_b}{n_t}\right)^{1/3},
\end{equation}
which connects with the condition on the reactive beam plasma instability from Equation (\ref{eq:bp reactive regime}).
Similarly for the ``oblique'' mode, the bandwidth, $\Delta \omega$, is then greater than the growth rate if 
\begin{equation}\label{eq:oblique kinetic regime}
\frac{\Delta v_{\perp}}{v_b} \gtrsim \left(\frac{n_b}{\gamma n_t}\right)^{1/3},
\end{equation}
which can similarly compared to the condition on the reactive ``oblique'' mode from Equation (\ref{eq:oblique reactive regime}).

Combining these two kinetic condition and our result again from Section \ref{sec:setup}, we find  
\begin{equation}\label{eq:generic kinetic regime}
1 \ge \left\{
\begin{array}{cl}
({k_BT}/{m_ec^2})^{-1/2}\gamma_b^{2/3}\left({n_b}/{n_t}\right)^{1/3} & \textrm{``oblique''} \\
({k_BT}/{m_ec^2})^{-1/2}\gamma_b\left({n_b}/{n_t}\right)^{1/3} & \textrm{beam plasma} 
\end{array}
\right.,
\end{equation}
which in combination with Equation (\ref{eq:generic reactive regime}) denotes the transition between the reactive and kinetic regimes.

@@

Still some progress can be made if we assume that the dominant loss rate is ICC so that we can place a lower limit on the various plasma processes.  In this case, the ratio of the beam plasma density to the IGM, $n_b/\nIGM$, is then (Paper I):
\begin{eqnarray}
\frac{\nb}{\nIGM} &\simeq&
\frac{L_E}{2\pi\Dpp^3\GIC}\frac{1}{\nIGM} \nonumber\\
&\simeq &
3\times10^{-18}
\left(\frac{1+z}{1}\right)^{3\zeta-7}
\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)
\left(\frac{E}{\TeV}\right)
\,\cm^{-3}\,,
\label{eq:nbIC}
\end{eqnarray}
where $\Dpp$ is defined in the Introduction to be the mean free path of a VHEGR, $L_E$ is the isotropic luminosity of the VHEGR source, $E$ is the energy of the VHEGR photon, $\Gamma_{\rm IC}$ is the inverse Compton cooling rate, and $\zeta = 4.5$ for $z \leq 1$ and $\zeta = 0$ for $z\geq 1$.

In Section \ref{sec:transition}, we derived the controlling parameter that delineates the reactive (eq.[\ref{eq:generic reactive regime}]) and kinetic regimes (eq.[\ref{eq:generic kinetic regime}]) by comparing the frequency spread of resonant waves, $\Delta\omega\sim k\Delta v$, with the growth rate, $\Gamma$.  Applying these conditions (eqns. [\ref{eq:generic reactive regime}] and [\ref{eq:generic kinetic regime}]) to the ultrarelativistic $e^+e^-$ pair beams of interest, we find for the controlling parameter: 
\begin{eqnarray}
\frac{\gamma_b^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2\times 10^{-2} \frac{\gamma_6^{2/3}}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam oblique condition}
\end{eqnarray}
and
\begin{eqnarray}
\frac{\gamma_b}{\sqrt{k_BT/m_e c^2}}\left(\frac{n_b}{\nIGM}\right)^{1/3} &=& 3.2 \frac{\gamma_6}{\sqrt{k_BT/m_e c^2}}\left(\frac{1+z}{2}\right)^{\zeta-7/3}\nonumber\\
&&\left(\frac{E L_E}{10^{45}\erg\,\s^{-1}}\right)^{1/3}
\left(\frac{E}{\TeV}\right)^{1/3},\label{eq:beam bp condition}
\end{eqnarray}
where $\gamma_6 = \gamma/10^6$. 
We see from our reactive (\ref{eq:generic reactive regime}) and kinetic (\ref{eq:generic kinetic regime}) conditions, that the ``oblique'' instability always exists in the kinetic regime, but the beam plasma instability sits in the reactive regime at $z\gtrsim 1$ or for sufficiently cold beams $k_BT/m_e c^2 \lesssim 0.5$, which occurs $E_{\rm ph} \sim 1\,\TeV$ and for large $\gamma$, which occurs for $E_{\rm ph} \sim 10\,\TeV$ at $z=0$.

The discussion above confirmed our earlier estimate in Paper I that the ``oblique'' instability sits in the kinetic regime.  However, our new results suggest that the beam-plasma instability matches the ``oblique'' instability in effective growth rate in the kinetic regime and that the beam-plasma mode may also sit in the reactive regime.  However, we do not expect that the reactive beam-plasma mode will have a major effect on our earlier results.  First, plasma instabilities losses on the TeV pairs could easily push the beam plasma mode into the kinetic regime by reducing $n_b$, but this requires a proper estimate of the effect of the nonlinear instability.  This is a part of ongoing work and will be presented in a future publication. Second, while it seems that the beam plasma mode may sit in the reactive regime, it is not too far from the kinetic regime, i.e., the controlling parameter, $({\gamma_b}/{\sqrt{k_BT/m_e c^2}})\left({n_b}/{\nIGM}\right)^{1/3}$, is order unity.  Thus, both the reactive and the kinetic growth rates are similar and likely makes little difference for the beam plasma mode which regime is assumed (in terms of growth rate).  Therefore, the use of the kinetic growth rate for the ``oblique'' mode (and beam-plasma mode)  in Paper I is valid, and the results of this paper buttresses the results of Paper I, II, III, and \citet{Puchwein+2011}.

@@

The ultrarelativistic $\epm$ beams that result from VHEGR-EBL annihilation are subject to powerful plasma beam instabilities including the beam plasma and ``oblique'' instability.  In this work, we have these linear instabilities as they would apply to the ultrarelativistic pair beams.  Our main findings are:
\begin{itemize}
%\item{We calculate how the parallel and perpendicular velocity dispersions of a distribution function responds under Lorentz boosts.  We show that these two velocities dispersions respectively scale like $\gamma^{-2}$ and $\gamma^{-1}$, allowing us to develop a simplified distribution function of the TeV electron-positron beam. We also show that this can be identified with a drifting relativistic Maxwell-Juttner distribution. }
\item{We analytically calculated the beam-plasma and ``oblique'' instabilities in both the reactive and kinetic regime.  We have recovered the reactive scalings for the beam-plasma mode $\Gamma \sim \gamma^{-1}(n_b/n_t)^{1/3}$ and the ``oblique'' mode $\Gamma \sim (n_b/\gamma n_t)^{1/3}$.  In the kinetic regime, we have shown that the growth rate for both modes have the same scaling and similar normalization.}
\item{We also delineated the regime of applicability of the kinetic and reactive calculation and found, while the kinetic growth rates are similar for both the beam plasma and ``oblique'' mode, the condition for transition between the kinetic and reactive regimes for both are different.  In particular, the beam-plasma mode transitions at a lower value of $\gamma$ in comparison to the ``oblique'' mode.  This is due to a difference of $\gamma^{1/3}$ scaling between the two modes.}
\item{We calculate the average COM energy of the ultrarelativistic pair beam using a simplified model of the spectra of the EBL.  We found that the average energy of these beams range from $E_e/m_e c^2 = 1.3-3$ for $E_{\rm ph} = 1-10\,\TeV$, with colder beams at lower energies.  This is due to the rapid cutoff in the EBL above $10\,\eV$. }
\item{We demonstrate that the ``oblique'' mode is in the kinetic regime, while, the beam plasma mode is marginally in the reactive regime. In any case, the use of either regime produces roughly the same growth rate.  Moreover, the inclusion of plasma instability losses on the number density of beam particles may place the beam plasma mode firmly in the kinetic regime.}
\end{itemize}

%The results of this paper are a first step toward an understanding of the nonlinear regime of these plasma instabilities for ultrarelativistic \epm\ beams.  In particular, by developing a simple distribution function, whose modes can be analytically calculated, we will develop a quasilinear theory for these modes in a forthcoming publication.  Such a quasilinear calculation will capture the dominant wave-particle interaction and capture the initial nonlinear evolution of this instability.

@@

\section{Lorentz Dependence of the Distribution Function and Velocity Dispersion}
\label{sec:lorentz}

Here we explicitly derive the scaling of the parallel and
perpendicular velocity dispersions with the Lorentz factor upon boosting the
distribution function to the lab frame. Let us begin with a distribution
function that is isotropic in the COM frame and depends only on energy.
Therefore in the COM frame, which we denote with the subscript ``CM'', the distribution
function is  $\fCM(E_{\rm CM}(\ppos_{\rm CM},\pmom_{\rm CM}))$.  When we move to the lab
(denoted with subscript ``L'') frame, the integral of the distribution function remains invariant,
i.e., total number, or
\begin{equation}
 \int \fLAB \dVLAB = \int \fCM \dVCM.
\end{equation}
It is well-known that under Lorentz transformations (Landau), 
\begin{equation}
 \dVLAB = \dVCM,
\end{equation}
so therefore,
\begin{equation}
  \fLAB(\ppos_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM}),\pmom_{\rm L}(\ppos_{\rm CM},\pmom_{\rm CM})) = \fCM(E_{\rm CM}).
\end{equation}

Now let us consider moments of the distribution function.  For clarity, it is helpful to consider moments of the distribution function first in the COM frame.  The velocity moment is (unsurprisingly):
\begin{equation}
 \betabarCM = \int \frac {\pmom_{\rm CM}}{\gamma_{\rm CM} mc} \fCM \dVCM = 0.
\end{equation}
We consider the lab frame to be  boosted along the x-axis by $\betaBOOST$. More precisely, the initial inertial frame is the COM frame and the lab frame is moving with velocity $\betaBOOST = -|\betaBOOST |$ with respect to the COM frame. This gives:
\begin{eqnarray}\label{eq:betabarLAB}
 \betabarLAB &=& \int \frac {\pmom_{\rm L}}{\gamma_{\rm L} mc} \fLAB \dVLAB  \nonumber\\
&=& \int \frac {\pmom_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM})}{\gamma_{\rm L}(\pmom_{\rm CM}, \ppos_{\rm CM}) mc} \fCM \dVCM.
\end{eqnarray}
Breaking the components of \betabarLAB\ into components parallel and
perpendicular to the boost, we find: 
\begin{eqnarray}
\betabarLABpara &=& \int \frac{\betaCMpara -\betaBOOST}{1-\betaBOOST\betaCMpara} \fCM \dVCM,\\
\betabarLABperp &=& \int \frac{\betaCMperp}{\gamma_{\rm B}\left(1-\betaBOOST\betaCMpara\right)} \fCM \dVCM,
\end{eqnarray}
where $\gamma_{\rm B} = \left(1-\betaBOOST^2\right)^{-1/2}$ is the Lorentz factor of the boost between the lab and COM frame.  
For $\left|\betaCM\right|,\betaBOOST \ll 1$, we recover the Galilean invariant result, $\betabarLAB \approx \betabarCM - \betaBOOST\hat{\boldsymbol x}$.  
However, this Galilean result no longer holds for relativistic motion.
